---
title: "Tutorial Edge and Nodelists"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial-Edge-Nodelists}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create edge and nodelists

## 1: Install `cognitivemapr`

To install the cognitivemapr package, run the following code:

```{r, results='hide', warning=FALSE, message=FALSE}
install.packages("devtools", repos = "https://cloud.r-project.org/")
library(devtools)

devtools::install_github("https://github.com/Fesch-star/cognitivemapr")

# install.packages("remotes", repos = "https://cloud.r-project.org/")
# library(remotes)

# remotes::install_github("Fesch-star/cognitivemapr")
```

## 2: Introducing the functions

The cognitivemapr package contains three functions that help to reformat
existing edgelists, create associated nodelists and put these into the
proper form to be used with the functions in this package. A cognitive
map is a semantic network consisting of causal and utility relations
(see Van Esch and Snellens, 2024). Therefore, in order to use this
Rpackage, the edgelist should consist of a list of directed relations
between concepts. The nodelist contains all unique concepts in the
edgelist with additional information regarding these concepts, such as
their intrinsic normative value or their categorisation in terms of
paradigms or (policy) instruments (see further below). The functions
that are introduced in this tutorial cater to the various startingpoints
the data may find itself in:

1.  prepare_lists_from_edgelist: This function is used when you only
    have an edgelist with relations. It helps you create the nodelist
    associated with this edgelist and reformats the edgelist to work
    with the other functions in the cognitivemapr package;

2.  align_edge_nodelist: This function is used when you have an edgelist
    and a nodelist, but these are not properly aligned yet. The function
    (re)aligns the lists by copying the (normative) value of nodes from
    the nodelist to the edgelist, which is required for running the
    other functions in the package. This situation may occur when you
    have (manually) added or changed the (intrinsic normative) value of
    some of the nodes in your nodelist (you do not need to re-align the
    edge and nodelist returned from the first function when you have
    only added the categorisation in terms of paradigms or instruments).

3.  reformat_edge_nodelist: This function is used when you have both an
    edgelist and associated nodelist that contain the necessary
    information but are not in the required format for this package. The
    function reformats both lists so you can run the other functions in
    this package.

The edgelists and nodelists that these functions take, already require
you to use a certain formatting while compiling them (see below). In
this tutorial simple bespoke edgelists (and nodelists) will be created
as examples to run the functions on.

## 3: Create a nodelist from an edgelist

In order to run the functions in this package an edgelist with
associated nodelist is needed. If your data only exists of an edgelist
with relations, the function "prepare_lists_from_edgelist" will help you
create an accompanying nodelist, containing all the (unique) concepts in
your edgelist.

This function requires your edgelist to have the following 4 columns
with the listed column titles, in addition any number of columns with
meta-data may be added.

1.  "from": cause concept (may consist of multiple words)

2.  "to": effect or utility concept (may consist of multiple words)

3.  "weight": number of times the relation is mentioned (can have any
    number, if left empty the function will set it to 1 by default)

4.  "edge_value": sign/value of the relation: can be 1 (positive
    relation), -1 (negative relation), or 0 (non-existent relation). If
    left empty the function will set it by default to 1 (positive).

We first create a simple dataframe/edgelist fitting these requirements
to use as an example in this tutorial (see Van Esch etal 2018 for a
discussion on how to derive an edgelist for CM analysis from texts or
obtain the data directly from respondents).

```{r message=FALSE, warning=FALSE, results='hide'}

# create a simple single edgelist

# create a list of cause concepts (mandatory column)
from <- c("a","c","d","b") 
# creat a list of effect/utility concepts (mandatory column)
to <- c("c","d","e","d") 
# set number of times the relation occurs (mandatory column, could be left empty)
weight <- c(1,1,1,1) 
# set the value of the relation (positive, negative or non-existent effect, could be left empty)
edge_value <- c(1,-1,-1,1) 
# example of non-mandatory meta-data: date
map_date <- c("20-5-2010", "21-5-2010","22-5-2010","23-5-2010") 
# combine the data into a dataframe: the original simple edgelist
single_edgelist <- data.frame(from, to, weight, edge_value, map_date)

```

Run the function on the example data and return the resulting edgelist
and nodelist.

```{r}
# run the function on the example data and return the resulting edge and nodelist

# run the function and return both the edge and nodelist
new_edgelist <- cognitivemapr::prepare_lists_from_edgelist(single_edgelist)[[1]]
nodelist <- cognitivemapr::prepare_lists_from_edgelist(single_edgelist)[[2]]

# show both lists
new_edgelist
nodelist


```

Reviewing the resulting edgelist and nodelist shows that the list has
been reformatted to fit the requirements of the other functions in this
package: The created nodelist contains all (unique) concepts featured in
the original edgelist and for each of these concepts a unique id was
created. By default, the function assumes that all concepts have a
positive/neutral (intrinsic normative) value (value of 1). As no
information was supplied about the categorisation of the concepts in
terms of paradigms or instruments, these columns are created but left
empty.

In addition, the edgelist has been reformatted to align with the
nodelist: Rather than listing the concepts, the "from" and "to" columns
now list the id's for the concepts corresponding to those in the
nodelist. The data in the columns "weight", "edge-value" and "map_date"
are left untouched. However two extra columns were added displaying the
(intrinsic normative) value of the concepts in the "from" (value.x) and
"to" (value.y) columns as derived from the nodelist.

The nodelist and new edgelist that the prepare_lists_from_edgelist
function creates contain the minimum amount of data needed to run the
most basic functions in the cognitivemapr Rpackage (calculate_degrees &
evaluate_concepts) as well as the correct formatting. However, more
information is needed to reap the full potential of the package, and to
use functions like paradigm_support, instrument_support and
interactive_visual_cm. These functions require a nodelist that supplies
the categorisation of concepts in terms of paradigms and instruments.
Also, some concepts may have an intrinsic negative normative value,
which necessitates adjusting the values in the "value" columns in both
the nodelist and edgelist. The nodelist that the
prepare_list_from_edgelist function has returned may be used as a
template to add/change this information.

## 4: Adding to or revising a nodelist

Deriving a nodelist from an edgelist with the prepare_list_from_edgelist
function results in nodelist that does not contain the necessary
categorisations of the concepts in the CM to conduct a full analysis. In
addition, the (normative) value of all concepts is set to
positive/neutral by default, which may not be accurate. In order to add
and/or change the information, changes must be made to the nodelist.
This may be done manually in excel by storing the nodelist. However, for
our example data, we will make the changes in R. When changes are made
in de value of the concepts in the nodelist, the edge and nodelist must
subsequently be realigned using the second list function in the
cognitivemapr package: align_edge_nodelist.

```{r}

# to store the new_nodelist when you want to make the changes manually in excel, you run the following code to save the list in csv format in a location of your choice [will not be executed in this tutorial]

# write.csv(nodelist, file = "[location/filename.csv]", row.names = FALSE)

# change nodelist by adding information on paradigms and instruments
nodelist$paradigms[nodelist$node_name == c("a", "b")] <- "ordoliberal" 
nodelist$instruments[nodelist$node_name == c("a")] <- "economic stimulation"

# change the value of concept a from positive to negative
nodelist$value[nodelist$node_name == "c"] <- -1

# renaming the nodelist for clarity
new_nodelist <- nodelist

# show the new_nodelist
new_nodelist
```

The new_nodelist that is returned has not changed and still contains the
same information as we (manually) added to it in the previous steps. The
new_edgelist2 shows a new value of -1 for concept c, which corresponds
to the changes we (manually) made to the nodelist. Note that there is no
trace of the changes we made in the paradigms and instruments column in
the edgelist. This information is only stored in the nodelist. This
means no re-alignment of the egdelist and nodelist is needed when you
only amend the information on the paradigms and instruments.

## 5: (Re) align the edge and nodelist

The second function to create edge and nodelist that are properly
formatted to use with the functions in the cognitivemapr Rpackage -
align_edge_nodelist - may be used to properly realign the edgelist with
the amended nodelist. This function takes the amended nodelist and the
edgelist which was returns from the 'prepare_lists_from_edgelist' as
it's input.

```{r}

#run the function on the amended nodelist and new_edgelist
new_edgelist2 <- cognitivemapr::align_edge_nodelist(new_edgelist, nodelist)[[1]]
new_nodelist <- cognitivemapr::align_edge_nodelist(new_edgelist, nodelist)[[2]]

# show the new edge and nodelist
new_edgelist2
new_nodelist
```

The new_nodelist that is returned has not changed since the previous
step and still contains the same information as we (manually) added to
it. The new_edgelist2, however, shows a new value of -1 for concept c,
which corresponds to the changes included in the new_nodelist. Note that
there is no trace of the changes we made in the paradigms and
instruments column in the edgelist. This information is only stored in
the nodelist. This means no re-alignment of the egdelist and nodelist is
needed when you only amend the information on the paradigms and
instruments.

## 6: Reformat edge and nodelist

Finally, you may start the proces of analysing your data with the
cognitivemapr package with both an edgelist as well as a nodelist with
the proper node values and categorisation of the concepts in terms of
paradigms and instruments. However, these may not yet fulfill the
formatting requirements of the package. The function
reformat_edge_nodelist helps you put a simple edge and nodelist in the
proper formatting to work with the other cognitivemapr functions. There
are some simple requirements for the function to reformat your lists.

The reformat_edge_nodelist function requires your **edgelist** to have
the following 4 columns with the listed column titles, in addition any
number of columns with meta-data may be added.

1.  "from": cause concept (may consist of multiple words)

2.  "to": effect or utility concept (may consist of multiple words)

3.  "weight": number of times the relation is mentioned (can have any
    number, if left empty the function will set it to 1 by default)

4.  "edge_value": sign/value of the relation: can be 1 (positive
    relation), -1 (negative relation), or 0 (non-existent relation). If
    left empty the function will set it by default to 1 (positive).

The reformat_edge_nodelist function also requires your **nodelist** to
have the following 4 columns with the listed column titles, in addition
any number of columns with meta-data may be added.

1.  "node_name": name of all unique concepts in the edgelist

2.  "value": the intrisic (normative) value of the concept displayed as
    a number (1 = positive/neutral, -1 = negative). You should use the
    negative value sparsely for whether a concept is positive or
    negative is often subjective. If a cell that is left empty in this
    column the function will set it by default to 1 (positive/neutral).

3.  "paradigms": the categorisation of the concept as belonging to/being
    associated with one of two incommensurable paradigms, such as
    Ordoliberal versus Keynesian (van Esch & Snellens 2024). You can
    only distinguish two paradigms in this column or leave the cell
    empty when a concept does not align with one of these two paradigms.

4.  "instruments": the categorisation of the concept as a particular
    type of policy-instrument. You can distinguish any number of
    different instruments in this column or leave the cell empty when a
    concept is not a policy instrument. This column can also be used for
    any other type of categorisation of the concepts that is deemed
    relevant, but you cannot change the name of the column.

The single_edgelist created at the start of this tutorial fulfills the
requirements for the reformat_edge_nodelist function, however, we do
need to create a new nodelist for the one's created by the previous
functions, already has more columns than required.

```{r}

# create a simple nodelist with the proper information on the value, paradigms and instruments  
node_name <- c("a","b","c","d", "e") # making a list of all unique concepts in the edgelist
value <- c("1","-1","1","1","-1") # making a list of node values
paradigms <- c("ordoliberal", NA, NA, "keynesian", NA) # list of paradigms
instruments <- c("economic stimulation", "financial market regulation", NA, NA, "economic stimulation") # list of instruments
date <- c("20-5-2010", "20-5-2010","22-5-2010","23-5-2010", "23-5-2010") # non-mandatory meta-data

# compile the nodelist (dataframe) from these lists
complete_nodelist <- data.frame(node_name, value, paradigms, instruments, date)

#show the complete_nodelist
complete_nodelist
```

We now have both a simple edge and nodelist that may easily be compiled
by the researcher and that has the proper format to run the
reformat_edge_nodelist function on.

```{r}
# run the function & return the edge and nodelist

new_edgelist3 <- cognitivemapr::reformat_edge_nodelist(single_edgelist, complete_nodelist) [[1]]
new_nodelist2 <- cognitivemapr::reformat_edge_nodelist(single_edgelist, complete_nodelist) [[2]]

new_edgelist3
new_nodelist2
```

Reviewing the returned edge and nodelist shows that the
reformat_edge_nodelist function adds id's to the nodelist and replaces
the concepts in the edgelist with the id's that correspond to those in
the nodelist. It reorders the columns in the edgelist to match the
requirements of the functions in this package. It checks whether the
columns weight and edge-value in the edgelist and the column value in
the nodelist have values, and sets these values to the default of 1 if
they were left empty in the uploaded lists. It assigns the "from" and
"to" concepts in the edgelist the value they were given in the nodelist
(or sets to the default of 1 if left empty) by adding respectively the
value.x and value.y to the edgelist. It The function also retains all
meta-data that was included in the original lists.

These list can now be used for analysis with the other functions in the
cognitivemapr package.

```{r}
#running the first function of the package - calculate_degrees - with the new_edgelist3 and new_nodelist2 as an example data of Mark Rutte, and storing it as a df
test_node_measures <- cognitivemapr::calculate_degrees(new_edgelist3, new_nodelist2)

# this shows that the created edge and nodelist indeed work with the other functions in the package and returns some first basis findings
str(test_node_measures)
```

Please refer to the [cognitivemapr Tutorial](cognitivemapr-Tutorial.html) for an explanation and
illustration of the analytical functions in the cognitivemapr package.
